---
title: "Ple.27.3a.21-32 Annual Assessment Model Interrogation"
author:
  - name: "Elliot J. Brown"
  - name: "Sven Stötera"
  - name: "Casper W. Berg"
date: "`r format(Sys.Date(), '%d/%m/%Y')`"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    code_folding: hide
    code_download: true
---

<style type="text/css">
.main-container {
  max-width: 90% !important;
  margin: auto;
}
p.caption {
  font-size: 0.8em;
  font-style: italic;
}
</style>

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=TRUE,
                      eval = TRUE,
                      warning = FALSE,
                      fig.align = "center",
                      tidy = 'formatR',
                      out.width = "90%")

#===
# Manual settings for figure formatting
#====
# ebpal <- c("#2F3EEA", "#1FD082", "#030F4F", "#F6D04D", "#FC7634", "#F7BBB1", "#E83F48", "#008835", "#79238E")
ebpal <- c("#800080", "#006400", "#D55E00", "#0072B2", "#F0E442", "#009E73", "#E69F00", "#56B4E9", "#CC79A7", "#5DA5DA", "#FF8000", "#89CFF0", "#A52A2A", "#77DD77", "#FFFACD")
#=====
```

# Introduction
## About this document
This interactive document downloads, refits, visualises and interrogates the assessment model from stockassessment.org.  The purpose is to provide an thorough quality control check on the model's performance, and to ensure the working group's (and any other interested parties') understanding of stock dynamics and model estimations. 

Data compilation is done externally, and this document deals only with the finalised input files. All models are build using the State-Space Assessment Model (SAM), and in this version of the document, all model input files and configurations are downloaded directly from the publicly available versions on [stockassessment.org](https://stockassessment.org).

The script that generates this document is one in a series which contribute to the production of advice for this plaice stock. To download the entire series, and associated directory structure, the project is available at GitHub: https://github.com/e-j-brown/Ple.27.21_32_Public.

Most of the code used to process data and create visualisations are hidden in this HTML.  To view each chunk, simply click on the button to the right of the relevant section.  To download the entire R-markdown (.rmd) file including all text and code chunks, use the _code_ button at the top of the page, with the dropdown arrow.

## Preparation
To run the scripts embedded in this document (if you would like to run the assessment yourself) you require the following packages and functions:

```{r message=FALSE, results="hide"}
require(data.table)
require(reshape2)
require(knitr)
require(scales)
require(bookdown)
require(ggplot2)
require(ggthemes)
require(stringr)
require(lattice)
require(plyr)
require(plotly)
require(icesAdvice)
require("stockassessment") # available from https://github.com/fishfollower/SAM
require(parallel)
require(sf)
require(ggforce)

myprop <- function(df, nomo = "nomo", NoAtLngt = "CANoAtLngt"){
  t <- sum(df$nomo)/sum(df$CANoAtLngt)
}

catchFrac <- function(x, nm, w, frac){
  F <- getF(x)
  Z <- F+nm
  N <- getN(x)
  C <- F/Z*(1-exp(-Z))*N
  return(sum(frac*w*C))
}

apply_rounding <- function(df) {
  # Iterate through each column in the dataframe
  df[] <- lapply(df, function(col) {
    if (is.numeric(col)) {
      # Apply rounding rules to numeric columns, skipping NAs
      return(sapply(col, function(x) {
        if (is.na(x)) {
          return(NA)
        } else if (x < 0.2) {
          return(round(x, 3))
        } else if (x >= 0.2 && x < 1) {
          return(round(x, 2))
        } else {
          return(round(x, 0))
        }
      }))
    } else {
      # Leave non-numeric columns unchanged
      return(col)
    }
  })
  return(df)
}

make_plus_group <- function(N, M, cutage) {
  ## N : matrix of numbers-at-age (rows = years, cols = ages)
  ## M : matrix of mean-at-age (same dimensions as N)
  ## cutage : numeric age at which plus group starts
  
  ## --- basic checks ---
  if (!is.matrix(N) || !is.matrix(M))
    stop("N and M must both be matrices")

  if (!all(dim(N) == dim(M)))
    stop("N and M must have identical dimensions")

  ages <- as.numeric(colnames(N))
  if (any(is.na(ages)))
    stop("Column names of N must be numeric ages")

  if (!cutage %in% ages)
    stop("cutage not found in column names")

  ## --- index sets ---
  GE <- which(ages >= cutage)   # ages in plus group
  E  <- which(ages == cutage)   # plus-group column
  low <- which(ages < cutage)   # ages kept as-is

  ## --- construct plus group numbers (additive) ---
  N_plus <- rowSums(N[, GE, drop = FALSE])

  ## --- weights for means (numbers-weighted) ---
  w <- N[, GE, drop = FALSE] / rowSums(N[, GE, drop = FALSE])

  ## --- construct plus group mean ---
  M_plus <- rowSums(M[, GE, drop = FALSE] * w)

  ## --- build reduced matrices ---
  N_out <- N[, c(low, E), drop = FALSE]
  M_out <- M[, c(low, E), drop = FALSE]

  N_out[, E] <- N_plus
  M_out[, E] <- M_plus

  return(list(
    N = N_out,
    M = M_out
  ))
}

```

We also need to set a few variables manually.  
The first is the *DataYear*, which is the year immediately preceeding the assessment year or the year for which the most recent data is available.  
The second is the catch we (ICES) advised as TAC for the last assessment *LastIcesAdvice*, in tonnes. __NOTE__ this value is taken from the advice sheet, to compare changes in advice, not changes in advice relative to actual Allowable Catches.
The third is whether this run is exploratory, or to be used to generate the final advice. This option can be used to determine our conclusions and in some cases, the number of iterations used in parameter estimations. 

We also utilise some reference points to contextualise the output estimates of the model and aid in our understanding of stock status, so we bring those in here, based on the 2024 Benchmark workshop results.

Finally, we set a few paths to be able to find and save stuff.

```{r prepFixedVariables}
DataYear <- 2025
LastIcesAdvice <- 16533 # Advised catch for areas 21-32 (from stock splitting table in advice)
isFinal <- FALSE
# isFinal <- TRUE

## Reference points
MSYBtrigger <- 13460
Bpa <- 13460
Blim <- 11119
SSB_5pcMSY <- 11251.37
Fpa <- 0.149
Fmsy <- 0.149
Fmsy_lower <- 0.137
Fmsy_upper <- 0.149
Fmsy_unconstrained <- 0.46

Flim <- 0.57  # Not based on constraint rule and not used in advice but just FYI

## repository paths
readPath <- paste0("../", (DataYear+1), "/Surveys/")
dataPath <- paste0("DataProcessed/Surveys/")
figPath <- paste0("Figures/")
```

# Baseline

The baseline model we're using to comare this year's model against, is the final model from last year's Benchmark workshop, using catch data up until 2023 and survey data up to Q1 2025. This baseline model is called [ple.27.21-32_WGBFAS_2025_SA]https://stockassessment.org/setStock.php?stock=ple.27.21-32_WGBFAS_2025_SA) and can be found on [stockassessment.org](https://stockassessment.org).

```{r BASEFitFromWeb}
set.seed(12345)

## Primary attempt at getting latest fit from web
# fit_BASE <- fitfromweb("ple.27.21-32_WGBFAS_2025_SA")

## Back-up attempt if server unavailbe, use offline copy of model.Rdata
fit_BASE <- NULL
if (file.exists("PrefitModels/BaseFit.RData")) {

  load("PrefitModels/BaseFit.RData")
  fit_BASE <- fit

} else if (file.exists("PrefitModels/CurrentFit.RData")) {

  load("PrefitModels/CurrentFit.RData")
  fit_BASE <- fit

} else {

  message(
    "No model objects available. Please download model.RData objects from ",
    "stockassessment.org, rename as BaseFit.RData and CurrentFit.RData, ",
    "respectively, in the directory PrefitModels"
  )
}
rm(fit)
```

```{r dataframeSummary_BASE, results='hide', fig.show='hide'}
#===
# Create dataframe of current year's fit for plotting
#====
asum_BASE <- as.data.frame(summary(fit_BASE))
asum_BASE$Year <- as.integer(row.names(summary(fit_BASE)))
ct_BASE <- as.data.frame(catchtable(fit_BASE, obs.show = TRUE))
ct_BASE$Year <- as.integer(rownames(ct_BASE))
rownames(ct_BASE) <- NULL
ct_BASE <- rbind(ct_BASE, data.frame(Estimate = NA, Low=NA, High=NA, sop.catch=NA, Year=as.integer(DataYear)))
asum_BASE <- merge(x = asum_BASE, y = ct_BASE, by = "Year")
colnames(asum_BASE) <- c("Year", "R_age1", "Rlow", "Rhigh", "SSB", "SSBlow", "SSBhigh", "Fbar", "Flow", "Fhigh", "CatchEst", "Catchlow", "Catchhigh", "CatchObs")
#=====
```

# Current Assessment Model

## Fit Model

This year's assessment model is based on the new Stock Annex procedure from the 2024 Benchmark workshop. This model is also available from [stockassessment.org](https://stockassessment.org), and is called [ple.27.21.32_WGBFAS_2026_SA](https://stockassessment.org/setStock.php?stock=ple.27.21.32_WGBFAS_2026_SA).

```{r 2025_SAFitFromWeb}
set.seed(12345)

# ## Primary attempt to download latest fit from stockassessment.org
# fit_CURRENT <- fitfromweb("ple.27.21.32_WGBFAS_2026_SA")

## Back-up attempt if server unavailbe, use offline copy of model.Rdata
fit_CURRENT <- NULL
if (file.exists("PrefitModels/CurrentFit.RData")) {

  load("PrefitModels/CurrentFit.RData")
  fit_CURRENT <- fit

} else {

  message(
    "No model object available. Please download model.RData objects from ",
    "stockassessment.org, rename as CurrentFit.RData, in the directory PrefitModels"
  )
}
rm(fit)
```

```{r dataframeSummary_CURRENT, results='hide', fig.show='hide'}
#===
# Create dataframe of current year's fit for plotting
#====
asum_CURRENT <- as.data.frame(summary(fit_CURRENT))
asum_CURRENT$Year <- as.integer(row.names(summary(fit_CURRENT)))
ct_CURRENT <- as.data.frame(catchtable(fit_CURRENT, obs.show = TRUE))
ct_CURRENT$Year <- as.integer(rownames(ct_CURRENT))
rownames(ct_CURRENT) <- NULL
ct_CURRENT <- rbind(ct_CURRENT, data.frame(Estimate = NA, Low=NA, High=NA, sop.catch=NA, Year=as.integer(DataYear+1)))
asum_CURRENT <- merge(x = asum_CURRENT, y = ct_CURRENT, by = "Year")
colnames(asum_CURRENT) <- c("Year", "R_age1", "Rlow", "Rhigh", "SSB", "SSBlow", "SSBhigh", "Fbar", "Flow", "Fhigh", "CatchEst", "Catchlow", "Catchhigh", "CatchObs")
#=====
```

# Interrogate Model Fits
## Fits to Data

When the model is run we can easily see from warnings/errors if there is a convergence issue.  However, we can also confirm this, explicitly:

- The final model gradient: `r fit_CURRENT$opt$evaluations[2]`
- That there is a positive definite hessian: `r all(eigen(fit_CURRENT$opt$he)$values >0)`

Furthermore, SAM does not utilise "bounds" when fitting the model, and therefore, as standard check of whether model parameters are approaching their bounds is irrelevant.

First we can look at how the model fits to observations.  We can do that with two methods, one is looking at the difference in the log observations and estimate on the log scale, the second is looking at the standardised residuals.
- logObs − logPred tells you how much larger or smaller, multiplicatively
- residuals(fit) tells you how surprising that deviation is, given the assumed error model

The table below describes the different interpretations of these.

```{r tab_compare_fitmetrics}
# Create the comparison table
comparison_table <- data.frame(
  Aspect = c(
    "Scale",
    "Meaning",
    "Accounts for observation variance (σ)",
    "Comparable across ages / fleets",
    "Used directly in likelihood estimation",
    "Best suited for"
  ),
  `logObs - logPred` = c(
    "Log scale",
    "Log ratio of observation to prediction",
    "No",
    "No",
    "No",
    "Exploratory and visual fit diagnostics"
  ),
  `residuals(fit)` = c(
    "Standardized (dimensionless)",
    "Likelihood-based residual",
    "Yes",
    "Yes",
    "Yes",
    "Model checking and statistical inference"
  ),
  check.names = FALSE
)

# Render as an HTML table
kable(
  comparison_table,
  format = "html",
  caption = "Conceptual comparison of log-scale deviations and model residuals"
)

```


```{r fig.cap="This year's, stock annex based assessment model's fit to catch data by age (values on the model link function scale).", fig.height=8}
## Extract data from model object
logObs_CURRENT <- split(fit_CURRENT$data$logobs, ceiling(seq_along(fit_CURRENT$data$logobs)/fit_CURRENT$data$maxAgePerFleet[1]))
logPred_CURRENT <- split(fit_CURRENT$rep$predObs, ceiling(seq_along(fit_CURRENT$rep$predObs)/fit_CURRENT$data$maxAgePerFleet[1]))

## Transform data to useable dataframes
### Initialize an empty data frame for observations
logObs_CURRENT_df <- data.frame(age = integer(),
                                         fleet = integer(),
                                         year = integer(),
                                         logObs = numeric())

### Populate the data frame for observations
for (i in seq_along(logObs_CURRENT)) {
  # Calculate the group and year
  year <- ((i - 1) %/% 3) + 1
  fleet <- ((i - 1) %% 3) + 1
  
  ### Create a temporary data frame and append to the main data frame for observations
  temp_df <- data.frame(age = 1:7,
                        fleet = fleet,
                        year = year+2001,
                        logObs = logObs_CURRENT[[i]])
  logObs_CURRENT_df <- rbind(logObs_CURRENT_df, temp_df)
}

### Initialize an empty data frame for predictions
logPred_CURRENT_df <- data.frame(age = integer(),
                                          fleet = integer(),
                                          year = integer(),
                                          logPred = numeric())

### Populate the data frame for predictions
for (i in seq_along(logPred_CURRENT)) {
  # Calculate the group and year
  year <- ((i - 1) %/% 3) + 1
  fleet <- ((i - 1) %% 3) + 1
  
  ### Create a temporary data frame and append to the main data frame for predictions
  temp_df <- data.frame(age = 1:7,
                        fleet = fleet,
                        year = year+2001,
                        logPred = logPred_CURRENT[[i]])
  logPred_CURRENT_df <- rbind(logPred_CURRENT_df, temp_df)
}

## Plot
logPred_CURRENT_df$age <- as.character(logPred_CURRENT_df$age)
logObs_CURRENT_df$age <- as.character(logObs_CURRENT_df$age)

ggplotly(ggplot()+
           geom_point(data = logObs_CURRENT_df[logObs_CURRENT_df$fleet == 1 & logObs_CURRENT_df$year != (DataYear+1), ],
                      mapping = aes(x = year,
                                    y = logObs,
                                    colour = age),
                      shape = 21) +
           geom_line(data = logPred_CURRENT_df[logPred_CURRENT_df$fleet == 1 & logPred_CURRENT_df$year != (DataYear+1), ],
                     mapping = aes(x = year,
                                   y = logPred,
                                   colour = age)) +
           facet_wrap(facets = "age") + #, scales = "free_y") +
           scale_color_manual(values = ebpal, guide = FALSE) +
           guides(colour=FALSE, shape=FALSE) +
           theme_few())

# fitplot(fit_CURRENT, fleets=1)
```


```{r fig.cap="This year's, stock annex based assessment model's fit to Q1 survey data by age (values on the model link function scale).", fig.height=8}
ggplotly(ggplot()+
           geom_point(data = logObs_CURRENT_df[logObs_CURRENT_df$fleet == 2 & logObs_CURRENT_df$year != (DataYear+1), ],
                      mapping = aes(x = year,
                                    y = logObs,
                                    colour = age),
                      shape = 21) +
           geom_line(data = logPred_CURRENT_df[logPred_CURRENT_df$fleet == 2 & logPred_CURRENT_df$year != (DataYear+1), ],
                     mapping = aes(x = year,
                                   y = logPred,
                                   colour = age)) +
           facet_wrap(facets = "age") + #, scales = "free_y") +
           scale_color_manual(values = ebpal, guide = FALSE) +
           guides(colour=FALSE, shape=FALSE) +
           theme_few())
# fitplot(fit_CURRENT, fleets=2)
```

```{r fig.cap="This year's, stock annex based assessment model's fit to Q3/4 survey data by age (values on the model link function scale).", fig.height=8}
ggplotly(ggplot()+
           geom_point(data = logObs_CURRENT_df[logObs_CURRENT_df$fleet == 3 & logObs_CURRENT_df$year != (DataYear+1), ],
                      mapping = aes(x = year,
                                    y = logObs,
                                    colour = age),
                      shape = 21) +
           geom_line(data = logPred_CURRENT_df[logPred_CURRENT_df$fleet == 3 & logPred_CURRENT_df$year != (DataYear+1), ],
                     mapping = aes(x = year,
                                   y = logPred,
                                   colour = age)) +
           facet_wrap(facets = "age") + #, scales = "free_y") +
           scale_color_manual(values = ebpal, guide = FALSE) +
           guides(colour=FALSE, shape=FALSE) +
           theme_few())

# fitplot(fit_CURRENT, fleets=3)
```

```{r fig.cap="Standard Devations by Fleet; medium discard survival estimates.", height=8}
sdplot(fit_CURRENT, marg = c(5,4,1,1))
```

## Residuals
```{r calculateResiduals_CURRENT, warning=FALSE, results='hide'}
set.seed(12345)
resid_CURRENT <- residuals(fit_CURRENT)
resid_CURRENT_df <- data.frame(year = resid_CURRENT$year,
                                        fleet = resid_CURRENT$fleet,
                                        age = resid_CURRENT$age,
                                        observation = resid_CURRENT$observation,
                                        mean = resid_CURRENT$mean,
                                        residual = resid_CURRENT$residual)
resid_CURRENT_df$fleetName <- ifelse(resid_CURRENT_df$fleet == 1, attributes(resid_CURRENT)$fleetNames[1], 
                                              ifelse(resid_CURRENT_df$fleet == 2, attributes(resid_CURRENT)$fleetNames[2], 
                                                     ifelse(resid_CURRENT_df$fleet == 3, attributes(resid_CURRENT)$fleetNames[3], NA)))

resid_CURRENT_df$fleetAltName <- ifelse(resid_CURRENT_df$fleet == 1, "Residual Fishing", 
                                                 ifelse(resid_CURRENT_df$fleet == 2, "Q1 Surveys", 
                                                        ifelse(resid_CURRENT_df$fleet == 3, "Q3/4 Surveys", NA)))
```

```{r fig.cap="Estimated correlations in residual variation between ages for each of the fishing fleet (top) and the two surveys (middle & bottom), from the Model with medium discard mortality."}

if(!all(fit_CURRENT$conf$obsCorStruct=="ID")){ 
  corplot(fit_CURRENT)			  
  # setcap("Estimated correlations", "Estimates correlations between age groups for each fleet")
  # stampit(fit)
} else {
  print("No correlation structure configured for residuals across age.")
}

```

```{r fig.cap="One observation ahead residuals for the three fleets (red/pink = observation lower than model estimate, blue = observation greater than or equal to model estimate, size = magnitude of residual).", fig.height=8}
ggplotly(ggplot(resid_CURRENT_df) +
           geom_point(mapping = aes(x = year,
                                    y = age,
                                    size = abs(residual),
                                    colour = residual >= 0),
                      alpha = 0.7) +
           facet_grid(rows = "fleetAltName") +
           scale_colour_manual(values = c("TRUE" = ebpal[8], "FALSE" = ebpal[9])) +
           scale_y_continuous(limits = c(-1, 9), breaks = c(1:9)) +
           guides(colour=FALSE) +
           theme_few())

```

## Jittering
We can also test to see if the model is converging on some local minimum (i.e. it's fitting to some solution close to initialising values that represents noise and not the global solution, that is the proper solution).  To do this, we add random noise to the initial parameter values to see if the model will converge on a different minimum in it's data-space. The easiest way to investigate this is to see if the model fits vary alot depending on the starting values:
```{r Jitter_CURRENT}
set.seed(12345)
jit_CURRENT <- jit(fit = fit_CURRENT)

mt <- as.data.frame(modeltable(jit_CURRENT))
mt$model <- rownames(mt)
rownames(mt) <- NULL

kable(x = mt,
      digits = 3,
      caption = "Measures of  fit for a series of model refits with jitter applied to the input parameters.")
```

# Model Estimates

Now that we've looked at how the model is fitting and performing, we can consider how the estimates of the status look and compare them to current reference points.

```{r recreateAsum, eval=FALSE}
#===
# Create dataframe of current year's fit for plotting
#====
asum_CURRENT <- as.data.frame(summary(fit_CURRENT))
asum_CURRENT$Year <- as.integer(row.names(summary(fit_CURRENT)))
ct_CURRENT <- as.data.frame(catchtable(fit_CURRENT, obs.show = TRUE))
ct_CURRENT$Year <- as.integer(rownames(ct_CURRENT))
rownames(ct_CURRENT) <- NULL
ct_CURRENT <- rbind(ct_CURRENT, data.frame(Estimate = NA, Low=NA, High=NA, sop.catch=NA, Year=as.integer(DataYear+1)))
asum_CURRENT <- merge(x = asum_CURRENT, y = ct_CURRENT, by = "Year")
colnames(asum_CURRENT) <- c("Year", "R_age1", "Rlow", "Rhigh", "SSB", "SSBlow", "SSBhigh", "Fbar", "Flow", "Fhigh", "CatchEst", "Catchlow", "Catchhigh", "CatchObs")
#=====
```


```{r fig.cap= "Estimated spawning stock biomass for ple.27.21-32 and 95% confidence intervals (tonnes). Purple is the baseline model, while green is the outcome of the current stock annex procedure with the latest data. Reference points MSY B-trigger (orange), Bpa (orange, dotted),  and Blim (darker orange) are the horizontal lines."}
pssb <- ggplot() +
  geom_line(data = asum_BASE,
            mapping = aes(x=Year,
                          y=SSB,
                          series="Baseline"),
            colour = ebpal[1]) +
  geom_ribbon(data = asum_BASE,
              mapping = aes(x=Year,
                            ymin=SSBlow,
                            ymax=SSBhigh),
              fill = ebpal[1],
              alpha = 0.3) +
  geom_line(data = asum_CURRENT,
            mapping = aes(x=Year,
                          y=SSB,
                          series="Current Year SA"),
            colour = ebpal[2]) +
  geom_ribbon(data = asum_CURRENT,
              mapping = aes(x=Year,
                            ymin=SSBlow,
                            ymax=SSBhigh),
              fill = ebpal[2],
              alpha = 0.3) +
  geom_hline(yintercept = Bpa,
             colour = ebpal[3],
             linetype = "dotted",
             aes(series="Bpa")) +
  geom_hline(yintercept = MSYBtrigger,
             colour = ebpal[11],
             mapping = aes(series="MSY B-trigger")) +
  geom_hline(yintercept = Blim,
             colour = ebpal[3],
             mapping = aes(series="Blim")) +
  theme_clean()

ggplotly(pssb, tooltip = c("series", "x", "y"))
```


```{r fig.cap= "Annual fishing mortality estimates for ple.27.21-32 ages 3-5 and point wise 95% confidence intervals are shown by line and shaded area.   Purple is the baseline model, while green is the outcome of the current stock annex procedure with the latest data. Reference points Fmsy (orange), Fpa (orange, dotted), and Flim (darker orange) are the horizontal lines."}
pfbar <- ggplot() +
  geom_line(data = asum_BASE,
            mapping = aes(x=Year,
                          y=Fbar,
                          series="Baseline"),
            colour = ebpal[1]) +
  geom_ribbon(data = asum_BASE,
              mapping = aes(x=Year,
                            ymin=Flow,
                            ymax=Fhigh,
                            series="Baseline"),
              fill = ebpal[1],
              alpha = 0.3) +
  geom_line(data = asum_CURRENT,
            mapping = aes(x=Year,
                          y=Fbar,
                          series="Current Year SA"),
            colour = ebpal[2]) +
  geom_ribbon(data = asum_CURRENT,
              mapping = aes(x=Year,
                            ymin=Flow,
                            ymax=Fhigh,
                            series="Current Year SA"),
              fill = ebpal[2],
              alpha = 0.3) +
  geom_hline(yintercept = Fpa,
             colour = ebpal[3],
             linetype = "dotted",
             aes(series="Fpa")) +
  geom_hline(yintercept = Fmsy,
             colour = ebpal[11],
             aes(series="Fmsy")) +
  geom_hline(yintercept = Flim,
             colour = ebpal[3],
             aes(series="Flim")) +
  theme_clean()

ggplotly(pfbar, tooltip = c("series", "x", "y"))
```

```{r fig.cap= "Annual recruitment estimates for ple.27.21-32 and point wise 95% confidence intervals are shown by line and shaded area (numbers). Purple is the baseline model, while green is the outcome of the current stock annex procedure with the latest data."}
prec <- ggplot() +
  geom_line(data = asum_BASE,
            mapping = aes(x=Year,
                          y=R_age1,
                          series="Baseline"),
            colour = ebpal[1]) +
  geom_ribbon(data = asum_BASE,
              mapping = aes(x=Year,
                            ymin=Rlow,
                            ymax=Rhigh,
                            series="Baseline"),
              fill = ebpal[1],
              alpha = 0.3) +
  geom_line(data = asum_CURRENT,
            mapping = aes(x=Year,
                          y=R_age1,
                          series="Current Year SA"),
            colour = ebpal[2]) +
  geom_ribbon(data = asum_CURRENT,
              mapping = aes(x=Year,
                            ymin=Rlow,
                            ymax=Rhigh,
                            series="Current Year SA"),
              fill = ebpal[2],
              alpha = 0.3) +
  theme_clean()

ggplotly(prec, tooltip = c("series", "x", "y"))
```

```{r fig.cap= "Annual catch estimates and 95% confidence intervals (line and shaded area, respectively) for ple.27.21-32 and point annual observations (points).  Purple is the baseline model, while green is the outcome of the current stock annex procedure with the latest data, and orange are the observations."}
pcat <- ggplot() +
  geom_line(data = asum_BASE[asum_BASE$Year != (DataYear), ],
            mapping = aes(x=Year,
                          y=CatchEst,
                          series="Baseline"),
            colour = ebpal[1]) +
  geom_ribbon(data = asum_BASE[asum_BASE$Year != (DataYear), ],
              mapping = aes(x=Year,
                            ymin=Catchlow,
                            ymax=Catchhigh,
                            series="Baseline"),
              fill = ebpal[1],
              alpha = 0.3) +
  geom_line(data = asum_CURRENT[asum_CURRENT$Year != (DataYear+1), ],
            mapping = aes(x=Year,
                          y=CatchEst,
                          series="Current Year SA"),
            colour = ebpal[2]) +
  geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != (DataYear+1), ],
              mapping = aes(x=Year,
                            ymin=Catchlow,
                            ymax=Catchhigh,
                            series="Current Year SA"),
              fill = ebpal[2],
              alpha = 0.3) +
  geom_point(data = asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ],
             mapping = aes(x=Year,
                           y=CatchObs,
                           series="Observed Catches"),
             shape = 3,
             colour = ebpal[3],
             fill = ebpal[3]) +
  theme_clean()

ggplotly(pcat, tooltip = c("series", "x", "y"))
```
## Model Estimates Table
Here we create a table that can be copied into the ices SAG system to generate standard graphs.

```{r sagTable}
#===
# table of Recruitment, SSB, TSB and F
#====
## Create table from component parts
tsb<-tsbtable(fit_CURRENT)
colnames(tsb)<-c("TSB","Low", "High")
tabsum <- as.data.frame(cbind(summary(fit_CURRENT), tsb))
tabsum$Year <- as.integer(rownames(tabsum))
rownames(tabsum) <- NULL

colnames(tabsum) <- c("R_age1", "RLow", "RHigh", "SSB", "SSBLow", "SSBHigh", "Fbar", "FLow", "FHigh", "TSB", "TSBLow", "TSBHigh", "Year")
tabsum <- tabsum[, c("Year", "RLow", "R_age1", "RHigh", "TSBLow", "TSB", "TSBHigh",  "SSBLow", "SSB","SSBHigh", "FLow", "Fbar", "FHigh")]

## Format data appropriately
tabsum[tabsum$Year == (DataYear+1), c("Fbar", "FLow", "FHigh")] <- NA
tabsum <- apply_rounding(tabsum)

kable(tabsum,
      caption = paste0("Estimated recruitment, spawning stock biomass (SSB), total stock biomass (TSB) and fishing mortality for ages 3-6 of ple.27.21-23.")) 
```
## Fishing Pressure at Age
```{r fig.cap="Partial F at age; base model"}
### Data Prep
F <- as.data.frame(faytable(fit_CURRENT))
F$Year <- rownames(F)
rownames(F) <- NULL
FLng <- melt(data = F, id.vars = "Year", variable.name = "Age", value.name = "F_atAge")
FLng$Year <- as.numeric(as.character(FLng$Year))

### Plot
ggplotly(ggplot() + geom_line(data = FLng, mapping = aes(x = Year, y = F_atAge, colour = Age)) +
    scale_color_manual(values = ebpal, ) + scale_x_continuous(breaks = unique(FLng$Year)) +
    theme_few() + theme(axis.text.x = element_text(angle = 45)))
```


## Stock-Recruitment Plots
In ~2020, this stock experienced an inflection in stock size development, from a steady recovery to a population explosion. In the period prior to 2020 the stock recruitment relationship appeared random, therefore, geometric means of recruitment over the whole time series were used for interim year estimates in short-term forecasts. Since 2020, the rapid development of the stock and subsequent development of the recruitment have provided data in the data space that may indicate a more typical stock-recruitment relationship. 

Below we visualise this relationship with the latest data.  The datapoint for `r DataYear` is shaded a different colour, as the recruitment estimate is based on only a single datapoint (single observation of Age1 fish in the Q1 surveys), compared to `r DataYear-1` which has three in-year datapoints (Q1 & Q3/4 surveys, and catches) as well as data from the assessment year (Age2 fish from`r DataYear` Q1 surveys). Going back in time, the datapoints all get one extra datapoint contributing to the recruitment estimate (as the model fit is influenced by all data).

```{r StockRecruitDataPrep eval="FALSE"}
### Data Prep
#### Create new Dataframe with radial fades
# number of concentric ellipses per point
K <- 50

# precompute ellipse radii
a <- (asum_CURRENT$SSBhigh - asum_CURRENT$SSBlow) / 2
b <- (asum_CURRENT$Rhigh - asum_CURRENT$Rlow) / 2

# repeat rows K times
df_cloud <- asum_CURRENT[rep(seq_len(nrow(asum_CURRENT)), each = K), ]

# scaling factor for each ellipse
scale <- rep(seq_len(K) / K, times = nrow(asum_CURRENT))

# assign scaled ellipse parameters
df_cloud$a <- rep(a, each = K) * scale
df_cloud$b <- rep(b, each = K) * scale

# transparency (strongest at centre)
alpha_max <- 0.05
df_cloud$alpha <- alpha_max * ((1 - scale)^2)

#### add exception for datayear
df_cloud$use <- ifelse(df_cloud$Year == (DataYear+1), "FALSE", "TRUE")
tdf <- df_cloud[!duplicated(df_cloud$Year), ]

#### Fit Beverton-Holt model to estimated recruitment
asum_recent <- asum_CURRENT[asum_CURRENT$Year >= (DataYear-5) & asum_CURRENT$Year != (DataYear+1), ]

# Whole time series
bh_fit <- nls(
  log(R_age1) ~ log((alpha * SSB) / (1 + beta * SSB)),
  data = asum_CURRENT,
  start = list(alpha = 1, beta = 1e-4),
  algorithm = "port",
  lower = c(alpha = 0, beta = 0)
)

# Recent (more productive) period, without latest highly uncertain recruitment year
bh_fit_rec <- nls(
  log(R_age1) ~ log((alpha * SSB) / (1 + beta * SSB)),
  data = asum_recent,
  start = list(alpha = 1, beta = 1e-4),
  algorithm = "port",
  lower = c(alpha = 0, beta = 0)
)

#### Create "smooth" curve for ggplot
# Whole time series
bh_curve <- data.frame(
  SSB = seq(min(asum_CURRENT$SSB),
            max(asum_CURRENT$SSB),
            length.out = (nrow(asum_CURRENT)*10))
)
bh_curve$R_age1 <- (unname(coef(bh_fit)["alpha"])*bh_curve$SSB)/(1+unname(coef(bh_fit)["beta"])*bh_curve$SSB)

# Recent (more productive) period, excluding highly variable last year
bh_curve_rec <- data.frame(
  SSB = seq(min(asum_recent$SSB),
            max(asum_recent$SSB),
            length.out = (nrow(asum_recent)*10))
)
bh_curve_rec$R_age1 <- (unname(coef(bh_fit_rec)["alpha"])*bh_curve_rec$SSB)/(1+unname(coef(bh_fit_rec)["beta"])*bh_curve_rec$SSB)

```

```{r StockRecruitDataPrepSSBLag}
### Data Prep
#### Create new dataframe with one year advanced recruitments
asum_adv <- asum_CURRENT
asum_adv$SSB <- c(NA, asum_CURRENT$SSB[-nrow(asum_CURRENT)])
asum_adv$SSBlow <- c(NA, asum_CURRENT$SSBlow[-nrow(asum_CURRENT)])
asum_adv$SSBhigh <- c(NA, asum_CURRENT$SSBhigh[-nrow(asum_CURRENT)])

#### Create new Dataframe with radial fades
# number of concentric ellipses per point
K <- 50

# precompute ellipse radii
a <- (asum_adv$SSBhigh - asum_adv$SSBlow) / 2
b <- (asum_adv$Rhigh - asum_adv$Rlow) / 2

# repeat rows K times
df_cloud <- asum_adv[rep(seq_len(nrow(asum_adv)), each = K), ]

# scaling factor for each ellipse
scale <- rep(seq_len(K) / K, times = nrow(asum_adv))

# assign scaled ellipse parameters
df_cloud$a <- rep(a, each = K) * scale
df_cloud$b <- rep(b, each = K) * scale

# transparency (strongest at centre)
alpha_max <- 0.05
df_cloud$alpha <- alpha_max * ((1 - scale)^2)

#### add exception for datayear
df_cloud$use <- ifelse(df_cloud$Year == (DataYear+1), "FALSE", "TRUE")
tdf <- df_cloud[!duplicated(df_cloud$Year), ]

#### Fit Beverton-Holt model to estimated recruitment
asum_recent <- asum_adv[asum_adv$Year >= (DataYear-5) & asum_adv$Year != (DataYear+1), ]

# Whole time series
bh_fit <- nls(
  log(R_age1) ~ log((alpha * SSB) / (1 + beta * SSB)),
  data = asum_adv,
  start = list(alpha = 1, beta = 1e-4),
  algorithm = "port",
  lower = c(alpha = 0, beta = 0)
)

# Recent (more productive) period, without latest highly uncertain recruitment year
bh_fit_rec <- nls(
  log(R_age1) ~ log((alpha * SSB) / (1 + beta * SSB)),
  data = asum_recent,
  start = list(alpha = 1, beta = 1e-4),
  algorithm = "port",
  lower = c(alpha = 0, beta = 0)
)

#### Create "smooth" curve for ggplot
# Whole time series
bh_curve <- data.frame(
  SSB = seq(min(asum_adv$SSB, na.rm = T),
            max(asum_adv$SSB, na.rm = T),
            length.out = (nrow(asum_adv)*10))
)
bh_curve$R_age1 <- (unname(coef(bh_fit)["alpha"])*bh_curve$SSB)/(1+unname(coef(bh_fit)["beta"])*bh_curve$SSB)

# Recent (more productive) period, excluding highly variable last year
bh_curve_rec <- data.frame(
  SSB = seq(min(asum_recent$SSB),
            max(asum_recent$SSB),
            length.out = (nrow(asum_recent)*10))
)
bh_curve_rec$R_age1 <- (unname(coef(bh_fit_rec)["alpha"])*bh_curve_rec$SSB)/(1+unname(coef(bh_fit_rec)["beta"])*bh_curve_rec$SSB)

#### Save models for later reference in forecasts
save(bh_fit, file = "DataProcessed/Bev-Holt_fullTimeSeries.Rdata")
save(bh_fit_rec, file = "DataProcessed/Bev_Holt_truncTimeSeries.Rdata")
```

```{r fig.cap=paste0("SSB-Recruitment relationship(???).  Year labels represent recruitment year, SSB is lagged by age at recruitment (1). The orange line is the chronological path, the black line is the Beverton-Holt SR model for the whole timeseries, and the blue line is the Beverton-Hold SR model without the latest year (R with large uncertainty, in purple) and only the six years before that (i.e. ", DataYear-5, ":", DataYear, ")."}
# 
# ### Plot interactive
# ggplotly(
# ggplot() +
#   geom_rect(data = tdf,
#             mapping = aes(xmin = SSBlow, xmax = SSBhigh, ymin = Rlow, ymax = Rhigh, fill = use),
#             alpha = 0.15) +
#   geom_point(
#     data = tdf,
#     aes(x = SSB, y = R_age1, colour = use),
#     size = 0.75
#   ) +
#   geom_path(data = tdf,
#             mapping = aes(x = SSB, y = R_age1),
#             colour = ebpal[3]) +
#   scale_alpha_identity() +
#   scale_fill_manual(values = ebpal, guide = "none") +
#   scale_color_manual(values = ebpal, guide = "none") +
#   theme_few()
# )

### Plot static but elipses
ggplot() +
  geom_ellipse(
    data = df_cloud,
    aes(x0 = SSB, y0 = R_age1, a = a, b = b, angle = 0, alpha = alpha, fill = use),
    colour = NA
  ) +
  geom_point(
    data = tdf,
    aes(x = SSB, y = R_age1, colour = use),
    size = 0.75
  ) +
  geom_path(data = tdf,
            mapping = aes(x = SSB, y = R_age1),
            colour = ebpal[3]) +
  geom_text(
    data = tdf,
    aes(x = SSB, y = R_age1, label = Year),
    colour = "black",
    nudge_x = 0.03 * diff(range(tdf$SSB, na.rm = TRUE)),
    nudge_y = 0.03 * diff(range(tdf$R_age1, na.rm = TRUE)),
    size = 2.5,
    show.legend = FALSE
  ) +
  geom_line(
    data = bh_curve,
    aes(x = SSB, y = R_age1),
    linewidth = 1.1,
    colour = "black"
  ) +
  geom_line(
    data = bh_curve_rec,
    aes(x = SSB, y = R_age1),
    linewidth = 1.1,
    colour = ebpal[4]
  ) +
  scale_alpha_identity() +
  scale_fill_manual(values = ebpal, guide = "none") +
  scale_color_manual(values = ebpal, guide = "none") +
  theme_few()
```

# Model Performance and Sensitivities
Now that we've interrogated the model's fit and interpreted its results, we should also see how dependent these results are on the input data. In the leave-one-out  and the retrospective analyses, we remove different tuning fleets and recent data, to see how robust the model's estimations are to the current context.

## Leave-One-Out Analyses
A leave-one-out analysis is a form of sensitivity analysis, showing the impact the data from each tuning fleet has on the estimation of the key variables being estimated; namely SSB, F and recruitment.  

First we must run the leave-one-out analysis which refits the model in two iterations, removing one survey at a time. Then we can plot each of these new model fits over the full model to see the impact the removal of each has. 

```{r LeaveOneOut_CURRENT}
set.seed(12345)
LO_CURRENT <- leaveout(fit_CURRENT)

#=== 
# Get data from sam objects and generate useable dataframes
#====
q1mat <- as.data.frame(summary(LO_CURRENT$`w.o. Q1IBTS+BITS1-7+`))
q3mat <- as.data.frame(summary(LO_CURRENT$`w.o. Q4IBTS+BITS1-7+`))


woq1 <- data.frame(Year = as.integer(row.names(summary(LO_CURRENT$`w.o. Q1IBTS+BITS1-7+`))),
                   SSB = q1mat$SSB,
                   Fbar = q1mat$`Fbar(3-6)`,
                   R_age1 = q1mat$`R(age 1)`,
                   CatchEst = catchtable(LO_CURRENT$`w.o. Q1IBTS+BITS1-7+`)[,1],
                   series = rep("wo_Q1", times = nrow(q1mat)))

woq34 <- data.frame(Year = as.integer(row.names(summary(LO_CURRENT$`w.o. Q4IBTS+BITS1-7+`))),
                    SSB = q3mat[, "SSB"],
                    Fbar = q3mat[,"Fbar(3-6)"],
                    R_age1 = q3mat[,"R(age 1)"],
                    CatchEst = c(catchtable(LO_CURRENT$`w.o. Q4IBTS+BITS1-7+`)[,1], setNames(NA, (DataYear+1))),
                    series = rep("wo_Q34", times = (nrow(q3mat))))


asum_CURRENT$series <- rep("full", times = nrow(asum_CURRENT))
asumi_CURRENT <- asum_CURRENT[, c("Year", "SSB", "Fbar", "R_age1", "CatchEst" ,"series")]
#=====

losum_CURRENT <- rbind(woq1, woq34, asumi_CURRENT)
losum_CURRENT[losum_CURRENT$Year == (DataYear+1), "Fbar"] <- NA  # Remove estimate for assessment year as it makes no sense to estimate without catch data.

```

```{r fig.cap="Leave-one-out re-fits for the model with medium discard survival (without Q1 survey = blue, without Q3/4 survey = purple), overlain with full model estimates (black line and grey ribbon) of SSB (top left), F (top right), Recruitment (bottom left), and catch (bottom right; observations as yellow + and ribbon broken by interactive tool)."}
# ssbplot(LO)
lossb_CURRENT <- ggplotly(ggplot() +
                                     geom_line(data = losum_CURRENT,
                                               mapping = aes(x=Year,
                                                             y=SSB,
                                                             colour = series)) +
                                     geom_ribbon(data = asum_CURRENT,
                                                 mapping = aes(x=Year,
                                                               ymin=SSBlow,
                                                               ymax=SSBhigh),
                                                 fill = "black",
                                                 alpha = 0.2) +
                                     scale_colour_manual(values = c("black", ebpal[8:9])) +
                                     theme_clean() +
                                     theme(axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95)))

# Fplot(LO)
lof_CURRENT <- ggplotly(ggplot() +
                                   geom_line(data = losum_CURRENT,
                                             mapping = aes(x=Year,
                                                           y=Fbar,
                                                           colour = series)) +
                                   geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != (DataYear+1),],
                                               mapping = aes(x=Year,
                                                             ymin=Flow,
                                                             ymax=Fhigh),
                                               fill = "black",
                                               alpha = 0.3) +
                                   scale_colour_manual(values = c("black", ebpal[8:9])) +
                                   theme_clean() +
                                   theme(axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95)))

# recplot(LO)
lorec_CURRENT <- ggplotly(ggplot() +
                                     geom_line(data = losum_CURRENT,
                                               mapping = aes(x=Year,
                                                             y=R_age1,
                                                             colour = series)) +
                                     geom_ribbon(data = asum_CURRENT,
                                                 mapping = aes(x=Year,
                                                               ymin=Rlow,
                                                               ymax=Rhigh),
                                                 fill = "black",
                                                 alpha = 0.3) +
                                     scale_colour_manual(values = c("black", ebpal[8:9])) +
                                     theme_clean() +
                                     theme(axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95)))

# Catch plot (LO)
loca_CURRENT <- ggplotly(ggplot() +
                                    geom_line(data = losum_CURRENT,
                                              mapping = aes(x=Year,
                                                            y=CatchEst,
                                                            colour = series))+
                                    geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != (DataYear+1), ],
                                                mapping = aes(x=Year,
                                                              ymin=Catchlow,
                                                              ymax=Catchhigh),
                                                fill = "black",
                                                alpha = 0.3) +
                                    geom_point(data = asum_CURRENT[asum_CURRENT$Year != (DataYear+1), ],
                                               mapping = aes(x=Year,
                                                             y=CatchObs),
                                               shape = 3,
                                               colour = ebpal[5],
                                               fill = ebpal[5]) +
                                    ylab("Catch (tonnes)") +
                                    scale_colour_manual(values = c("black", ebpal[8:9]))+ 
                                    theme_clean() +
                                    theme(axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95)))

layout(subplot(lossb_CURRENT, lof_CURRENT, lorec_CURRENT, loca_CURRENT, nrows = 2, shareX = TRUE, titleY = TRUE), showlegend=FALSE)
```

## Retrospectives
Retrospective analyses, iterate through removing the most recent years of data and refiting the model.  The results illustrate how much the current data is impacting upon how we view the past status of the stock. We do this for 5 iterations ( _peels_), removing up to five years of data.

The mohn's rho is a relative error statistic used to describe the difference between an estimate ( _SSB_, _F_, etc.) with a full time series compared to the truncated series.  Generally, in ICES assessments, Mohn's rhos below 0.2 and above -0.15 are generally acceptable.  

In the case of this stock, our assessment model is fit to Q1 `r DataYear+1` survey data, and so it makes estimates right up to the assessment year.  However, these estimates of stock size and attributes in the assessment year lack the input of the commercial catch data and the second tuning fleet. Therefore, before starting the retrospective analyses, we remove the `r DataYear+1` estimates and begin the five peels from `r DataYear`.

```{r calculateRetros_CURRENT}
set.seed(12345)
RETRO_CURRENT<-retro(fit_CURRENT, year=5)
set.seed(12345)
rho_CURRENT <- mohn(RETRO_CURRENT, lag = 1)

## Make RETROs in better plotting format and remove latest year with only one datapoint.
ret_CURRENT_df <- asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ]

for(i in 1:length(RETRO_CURRENT)){
  tsum <- as.data.frame(summary(RETRO_CURRENT[[i]]))
  tsum$Year <- as.integer(row.names(summary(RETRO_CURRENT[[i]])))
  tsum <- cbind(tsum, catchtable(RETRO_CURRENT[[i]], obs.show = TRUE))
  tsum$series <- as.character(rep(i, nrow(tsum)))
  colnames(tsum) <- c("R_age1", "Rlow", "Rhigh", "SSB", "SSBlow", "SSBhigh", "Fbar", "Flow", "Fhigh", "Year", "CatchEst", "Catchlow", "Catchhigh", "CatchObs", "series")
  tsum <- tsum[tsum$Year != max(tsum$Year), ]
  ret_CURRENT_df <- rbind(ret_CURRENT_df, tsum)
}
ret_CURRENT_df$series <- factor(ret_CURRENT_df$series, levels = c("full", "1", "2", "3", "4", "5"))
```

```{r fig.cap="Retrospective analyses for SSB (top-left), F for ages 3-5 (top-right), recruitment (bottom-left), and catch (bottom-right), from the model with medium discard survival."}
# ssbplot(RETRO)
retssb_CURRENT <- layout(ggplotly(ggplot() +
                                             geom_line(data = ret_CURRENT_df,
                                                       mapping = aes(x=Year,
                                                                     y=SSB,
                                                                     colour = series)) +
                                             geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ],
                                                         mapping = aes(x=Year,
                                                                       ymin=SSBlow,
                                                                       ymax=SSBhigh),
                                                         fill = "black",
                                                         alpha = 0.2) +
                                             annotate(geom = "text",
                                                      y = max(ret_CURRENT_df$SSBhigh)*0.85,
                                                      x = ((max(ret_CURRENT_df$Year)-min(ret_CURRENT_df$Year))*0.20)+min(ret_CURRENT_df$Year),
                                                      label = paste0("Mohn's Rho = ", round(rho_CURRENT[2], digits = 3))) +
                                             scale_colour_manual(values = c("black", ebpal[(length(ebpal)-4):length(ebpal)])) +
                                             theme_clean() +
                                             theme(legend.position = "none",
                                                   axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95))),
                                  showlegend = FALSE)

# Fplot(RETRO)
retf_CURRENT <- layout(ggplotly(ggplot() +
                                           geom_line(data = ret_CURRENT_df,
                                                     mapping = aes(x=Year,
                                                                   y=Fbar,
                                                                   colour = series)) +
                                           geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ],
                                                       mapping = aes(x=Year,
                                                                     ymin=Flow,
                                                                     ymax=Fhigh),
                                                       fill = "black",
                                                       alpha = 0.3) +
                                           annotate(geom = "text",
                                                    y = max(ret_CURRENT_df$Fhigh)*0.85,
                                                    x = ((max(ret_CURRENT_df$Year)-min(ret_CURRENT_df$Year))*0.80)+min(ret_CURRENT_df$Year),
                                                    label = paste0("Mohn's Rho = ", round(rho_CURRENT[3], digits = 3))) +
                                           scale_colour_manual(values = c("black", ebpal[(length(ebpal)-4):length(ebpal)])) +
                                           theme_clean() +
                                           theme(legend.position = "none",
                                                 axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95))),
                                showlegend = FALSE)

# recplot(RETRO)
retrec_CURRENT <- layout(ggplotly(ggplot() +
                                             geom_line(data = ret_CURRENT_df,
                                                       mapping = aes(x=Year,
                                                                     y=R_age1,
                                                                     colour = series)) +
                                             geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ],
                                                         mapping = aes(x=Year,
                                                                       ymin=Rlow,
                                                                       ymax=Rhigh),
                                                         fill = "black",
                                                         alpha = 0.3) +
                                             annotate(geom = "text",
                                                      y= max(ret_CURRENT_df$Rhigh)*0.85,
                                                      x = ((max(ret_CURRENT_df$Year)-min(ret_CURRENT_df$Year))*0.20)+min(ret_CURRENT_df$Year),
                                                      label = paste0("Mohn's Rho = ", round(rho_CURRENT[1], digits = 3))) +
                                             scale_colour_manual(values = c("black", ebpal[(length(ebpal)-4):length(ebpal)])) +
                                             theme_clean() +
                                             theme(legend.position = "none",
                                                   axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95))),
                                  showlegend = FALSE)

## Catch plot (RETRO)
retca_CURRENT <- ggplotly(ggplot() +
                                     geom_line(data = ret_CURRENT_df,
                                               mapping = aes(x=Year,
                                                             y=CatchEst,
                                                             colour = series))+
                                     geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ],
                                                 mapping = aes(x=Year,
                                                               ymin=Catchlow,
                                                               ymax=Catchhigh),
                                                 fill = "black",
                                                 alpha = 0.3) +
                                     geom_point(data = asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ],
                                                mapping = aes(x=Year,
                                                              y=CatchObs),
                                                shape = 3,
                                                colour = ebpal[5],
                                                fill = ebpal[5]) +
                                     ylab("Catch (tonnes)") +
                                     scale_colour_manual(values = c("black", ebpal[(length(ebpal)-4):length(ebpal)]))+ 
                                     theme_clean() +
                                     theme(axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95)))

ret_fp <- style(subplot(retssb_CURRENT, retf_CURRENT, retrec_CURRENT, retca_CURRENT, nrows = 2, shareX = FALSE, shareY = FALSE, titleY = TRUE),
                showlegend=FALSE,
                traces = c(8:((7*4)+2)))

layout(ret_fp, legend = list(orientation = 'h', x = 0.5, xanchor = 'center', y = -0.05, yanchor = 'top', borderwidth = 0))
```

# Catch Estimates With Surviving Discards

In our assessment procedure, we utilise published survival rates of discarded fish across various gears, seasons (Quarters), and management regions to estimate annual proportions of discarded fish that would survive. This fraction of "surviving discards", are removed from the catch data that is put into the assessment (e.g. canum.dat and lf.dat).

The assessment model is then run with only landings and "dead discards", contributing to total catch.  Thus the forecasts made above, also only include landings and "dead discards".  To re-introduce the surviving component of the catch, we assume that the annual proportion of surviving discards by age remains _status quo_, that is the forecast advice year  (`r DataYear+2`) has the same ratios of surviving discards as the input data year (`r DataYear`).

```{r catchEstSurviving, eval=TRUE}
# Surviving fractions of discards ----
# ## Matrix from csv
# survfrac <- read.csv("DataIn/CatchData/ple.27.21-32_DiscardSurvivalRates_2002_DataYear.csv")
# survfrac <- as.matrix(survfrac[, 2:ncol(survfrac)])
# rownames(survfrac) <- 2002:DataYear
# colnames(survfrac) <- 1:10

## Direct from DataCoordination Lowestoft file
survfrac <- read.ices("Assessment Input/StockAnnex/df_frac.dat")

# Cutdown ages of input data to 7+ group ----
catch.no<-read.ices("Assessment Input/StockAnnex/cn.dat")/1000
cutage<-7
low<-1
GE<-which(as.numeric(colnames(catch.no))>=cutage)
E<-which(as.numeric(colnames(catch.no))==cutage)
w<-catch.no[,GE]/rowSums(catch.no[,GE])
wex<-rbind(w,w[nrow(w),])

survfrac[,E]<-rowSums(survfrac[,GE]*w)
survfrac<-survfrac[,low:E]

# # Extract stock numbers at age from model
# sn_CURRENT <- ntable(fit_CURRENT)[!rownames(sn_CURRENT) %in% (DataYear+1), ]

# Get estimated Dead catch numbers at age from model ----
## Landings and DEAD Discards combined
dcn_CURRENT <- caytable(fit_CURRENT)[!rownames(caytable(fit_CURRENT)) %in% (DataYear+1), ]
## DEAD discard numbers at age from model
ddcn_CURRENT <- dcn_CURRENT*(1-fit_CURRENT$data$landFrac[, , "Residual catch"])
## TOTAL discard numbers at age
adcn_CURRENT <- ddcn_CURRENT/(1-survfrac)
## SURVIVING discard numbers at age
sdcn_CURRENT <- adcn_CURRENT*survfrac

# Add SURVIVING discards back to DEAD catch numbers at age from the model
cn_CURRENT <- dcn_CURRENT+sdcn_CURRENT

# Calculate catches (tonnes) by age using mean weights over time
## Dead discards
ddcaton_CURRENT <- ddcn_CURRENT*fit_CURRENT$data$disMeanWeight[, , "Residual catch"]
## All Discards
adcaton_CURRENT <- adcn_CURRENT*fit_CURRENT$data$disMeanWeight[, , "Residual catch"]
## Landings only
lcaton_CURRENT <- (dcn_CURRENT-ddcn_CURRENT)*fit_CURRENT$data$landMeanWeight[, , "Residual catch"]
## Totals with only DEAD discards
dtcaton_CURRENT <- ddcaton_CURRENT+lcaton_CURRENT
## Totals with ALL Discards
caton_CURRENT <- adcaton_CURRENT+lcaton_CURRENT
caton_CURRENT_Check <- cn_CURRENT*fit_CURRENT$data$catchMeanWeight[, , "Residual catch"]

# newnums <- rbind(dcn_CURRENT[nrow(dcn_CURRENT), ],
#       ddcn_CURRENT[nrow(ddcn_CURRENT), ],
#       adcn_CURRENT[nrow(adcn_CURRENT), ],
#       sdcn_CURRENT[nrow(sdcn_CURRENT), ],
#       cn_CURRENT[nrow(cn_CURRENT), ],
#       # rep(" - ", times=ncol(dcn_CURRENT)),
#       ddcaton_CURRENT[nrow(ddcaton_CURRENT), ],
#       adcaton_CURRENT[nrow(adcaton_CURRENT), ],
#       lcaton_CURRENT[nrow(lcaton_CURRENT), ],
#       dtcaton_CURRENT[nrow(dtcaton_CURRENT), ],
#       caton_CURRENT[nrow(caton_CURRENT), ],
#       caton_CURRENT_Check[nrow(caton_CURRENT_Check), ])
# newnums <- round(newnums, 3)
# rownames(newnums) <- c("CatchAgeTable", "DeadDiscards", "AllDiscards", "SurvivingDiscards", "TotalCatchNo", "DeadDiscardCATONDW", "AllDiscardCATONDW", "LandingsCATONLW", "TotalDeadCATONDwLw", "TotalCATONDwLw", "TotalCATONCw")

# Check totals align with catchTable ----
## Extract estimates from catchTable
fit_summary <- catchtable(fit_CURRENT)

## Total DEAD catch mass, all ages summed, from calcs using discards and landings independently
dlCatchEst <- rowSums(dtcaton_CURRENT)
## Total DEAD & SURVIVING catch mass, all ages summed, from calcs using discards and landings independently
sdlCatchEst <- rowSums(caton_CURRENT)
## Surviving catch mass all ages summed from calcs using discards and landings independently
sCatchEst <- sdlCatchEst - dlCatchEst

## Total dead catch mass, all ages summed, from cayTable*cw
CatchEst <- rowSums(dcn_CURRENT*fit_CURRENT$data$catchMeanWeight[, , "Residual catch"])


# ## totals calculated from total catch, from cayTable*cw
# data.frame(ManualCalcs=CatchEst, CatchTableVals=fit_summary[, "Estimate"], Difference=CatchEst-fit_summary[, "Estimate"])
# 
# ## totals from independent discard and landings calculations
# data.frame(ManualCalcs=dlCatchEst, CatchTableVals=fit_summary[, "Estimate"], Difference=dlCatchEst-fit_summary[, "Estimate"])
# 
# ## totals with re-added surviving discards
# data.frame(ManualCalcs=sdlCatchEst, CatchTableVals=fit_summary[, "Estimate"], Difference=sdlCatchEst-fit_summary[, "Estimate"], survivingDiscards=rowSums(sdcn_CURRENT*fit_CURRENT$data$disMeanWeight[, , "Residual catch"]))


# Create new ASUM with survivors added back ----
## Initialise new ASUM
asum_Survivors <- asum_CURRENT[asum_CURRENT$Year != (DataYear+1), ]
asum_Survivors$CatchSurv <- unname(sdlCatchEst)

## Pull in and clean full catch data
catch.no<-read.ices('Assessment Input/StockAnnex/cn.dat')/1000
catch.mean.weight<-read.ices('Assessment Input/StockAnnex/cw.dat')

catch.no[,E]<-rowSums(catch.no[,GE])
catch.no<-catch.no[,low:E]
catch.mean.weight[,E]<-rowSums(catch.mean.weight[,GE]*w)
catch.mean.weight<-catch.mean.weight[,low:E]

CatchData_NoSurvivals <- catch.no*catch.mean.weight

## Add total catch observations back to new ASUM
asum_Survivors$CatchObsSurv <- rowSums(CatchData_NoSurvivals)

```

## Catch plot with added survivors
```{r fig.cap= "Annual catch estimates (lines), 95% confidence intervals (ribbon) and point annual observations (points) for ple.27.21-32.  Green is the outcome of the current stock annex procedure with the latest data, orange are the observations the model is fit to (both without surviving discards), and blue are the values with suriving discards added back to the estimates."}
pcat <- ggplot() +
  geom_line(data = asum_Survivors,
            mapping = aes(x=Year,
                          y=CatchSurv,
                          series="Estimated Catch with Surviving Discards"),
            colour = ebpal[8],
            size = 1) +
  geom_point(data = asum_Survivors,
             mapping = aes(x=Year,
                           y=CatchObsSurv,
                           series="Observed Catches with Surviving Discards"),
             shape = 16,
             colour = ebpal[10],
             fill = ebpal[10]) +
  geom_line(data = asum_Survivors,
            mapping = aes(x=Year,
                          y=CatchEst,
                          series="Model Estimated Dead Catches"),
            colour = ebpal[2]) +
  geom_ribbon(data = asum_Survivors,
              mapping = aes(x=Year,
                            ymin=Catchlow,
                            ymax=Catchhigh,
                            series="Model Estimated Dead Catches"),
              fill = ebpal[2],
              alpha = 0.3) +
  geom_point(data = asum_Survivors,
             mapping = aes(x=Year,
                           y=CatchObs,
                           series="Observed Catches NO Surviving Discards"),
             shape = 3,
             colour = ebpal[3],
             fill = ebpal[3]) +
  theme_clean()

ggplotly(pcat, tooltip = c("series", "x", "y"))
```

# SSB broken down by desired and undesired components.
```{r fig.cap="Proportion of SSB that are above the midpoint between the mean landing weight and the mean discard weight, by age, over time"}
#===
# Data wrangling
#===
## Create threshold weight by age for each year
cn <- read.ices("Assessment Input/StockAnnex/cn.dat") # Total catch numbers
lf <- read.ices("Assessment Input/StockAnnex/lf.dat") # Landings catch numbers
df <- cn-lf                                           # Discard catch numbers
lw <- read.ices("Assessment Input/StockAnnex/lw.dat") # Mean individual weight in the landings
dw <- read.ices("Assessment Input/StockAnnex/dw.dat") # Mean individual weigh in the discards

lw <- make_plus_group(N = lf, M = lw, cutage = 7)
dw <- make_plus_group(N = df, M = dw, cutage = 7)
w_thr <- ((ifelse(lw$M-dw$M < 0, 0, lw$M-dw$M))/2)+dw$M  # Midpoint between dw and lw, made up by Elliot.  Could be some other relationship.

## Assume DataYear's preferred weight threshold for assessment year.
w_thr <- rbind(w_thr, w_thr[nrow(w_thr), , drop = FALSE])
rownames(w_thr)[nrow(w_thr)] <- as.character(as.integer(rownames(w_thr)[nrow(w_thr)]) + 1)

## Extract other necessary data from model
N <- t(exp(fit_CURRENT$pl$logN))    # array: age × year
W <- fit_CURRENT$data$stockMeanWeight  # matrix: age × year

# ## Guestimate CV for weights
# cv_w <- 0.2
# sigma <- sqrt(log(1 + cv_w^2))

#===
# Calculate CV for weights
#===
ca_hh_fin_bits_ibts <- readRDS(file = "DataProcessed/Surveys/Ple.27.21-32_ca-hh.RDS")
ca_q1<-subset(ca_hh_fin_bits_ibts,Quarter==1&Age!=-9&Age!=0&Age<8)

ca_q1$groupID <- interaction(ca_q1$Year,
                             ca_q1$Age,
                             drop = TRUE)
mean_w <- tapply(ca_q1$IndWgt, ca_q1$groupID , mean)
sd_w   <- tapply(ca_q1$IndWgt, ca_q1$groupID , sd)
n_w    <- tapply(ca_q1$IndWgt, ca_q1$groupID , length)
cv_w <- sd_w / mean_w

### set results to dataframe
cv_w_df <- data.frame(
  year = as.integer(sub("\\..*$", "", names(mean_w))),
  age  = as.integer(sub("^.*\\.", "", names(mean_w))),
  mean_weight = as.numeric(mean_w),
  sd_weight   = as.numeric(sd_w),
  cv_w        = as.numeric(cv_w),
  n           = as.integer(n_w)
)

## Get unique ages
ages <- sort(unique(cv_w_df$age))

## Pre-allocate vectors
cv_w_age <- numeric(length(ages))
n_total  <- numeric(length(ages))

## Loop over ages
for (i in seq_along(ages)) {

  a <- ages[i]

  ## Subset rows for this age
  idx <- cv_w_df$age == a

  cv_vals <- cv_w_df$cv_w[idx]
  n_vals  <- cv_w_df$n[idx]

  ## Exclude NA CVs (and corresponding n)
  ok <- !is.na(cv_vals)

  if (any(ok)) {
    cv_w_age[i] <- sum(n_vals[ok] * cv_vals[ok]) /
                   sum(n_vals[ok])
    n_total[i]  <- sum(n_vals[ok])
  } else {
    cv_w_age[i] <- NA
    n_total[i]  <- 0
  }
}

## Assemble final data frame
cv_by_age <- data.frame(
  age     = ages,
  cv_w    = cv_w_age,
  n_total = n_total
)

sigma <- sqrt(log(1 + cv_by_age$cv_w^2))
#===

#===
# Use mean and CV to calculate proportions of biomass
#===
## Convert mean to lognormal mu
mu <- log(W) - 0.5 * sigma^2

## Proportion of biomass above threshold
P_biom_above <- 1 - pnorm(
  (log(w_thr) - mu - sigma^2) / sigma  # mu-sigma^2 because estimating for biomass remove sigma if estimating for number????
)

## Proportion for all ages, weighted by catch numbers at age
### Match column and row numbers
rownames(N) <- rownames(P_biom_above)
colnames(N) <- colnames(P_biom_above)
TBay <- N*W

### Make overall annual proportion of biomass above threshold
asum_CURRENT$P_biom_above <- c(unname(make_plus_group(N = TBay, M = P_biom_above, cutage = 1)$M))

## Split SSB by proportions
asum_CURRENT$Desired_SSB <- asum_CURRENT$SSB*asum_CURRENT$P_biom_above
asum_CURRENT$Undesired_SSB <- asum_CURRENT$SSB*(1-asum_CURRENT$P_biom_above)

```

## SSB plot split by desired and undesired proportions
```{r fig.cap= "Estimated spawning stock biomass for ple.27.21-32 and 95% confidence intervals (tonnes). Purple is the baseline model, while green is the outcome of the current stock annex procedure with the latest data. Reference points MSY B-trigger (orange), Bpa (orange, dotted),  and Blim (darker orange) are the horizontal lines."}
## Data reshape
ssbd_df <- data.frame(
  Year  = rep(asum_CURRENT$Year, 2),
  SSB   = c(asum_CURRENT$Desired_SSB,
            asum_CURRENT$Undesired_SSB),
  SSBlow = rep(asum_CURRENT$SSBlow, 2),
  SSBhigh = rep(asum_CURRENT$SSBhigh, 2),
  Type  = rep(c("Desired SSB", "Undesired SSB"),
              each = nrow(asum_CURRENT)),
  Proportion = c(asum_CURRENT$P_biom_above, (1-asum_CURRENT$P_biom_above))
)

pssbD <- ggplot() +
  geom_col(data = ssbd_df,
            mapping = aes(x=Year,
                          y=SSB,
                          series="Current Year SA",
                          text = round(Proportion, 3),
                          fill = Type)) +
  geom_errorbar(data = asum_CURRENT,
              mapping = aes(x=Year,
                            ymin=SSBlow,
                            ymax=SSBhigh),
              colour = ebpal[3],
              alpha = 0.5) +
  geom_hline(yintercept = MSYBtrigger,
             colour = ebpal[11],
             mapping = aes(series="MSY B-trigger")) +
  geom_hline(yintercept = Blim,
             colour = ebpal[3],
             mapping = aes(series="Blim")) +
  scale_fill_manual(values = ebpal, guide = "none") +
  theme_clean()

ggplotly(pssbD, tooltip = c("series", "x", "y", "fill", "text"))
```

## Summary of the assessment table

```{r assessmentSummarySurvivors}
## Assemble component parts
sumassess <- asum_CURRENT[, !grepl("Catch", colnames(asum_CURRENT))]
sumassess[sumassess$Year == (DataYear+1), grepl("F", colnames(sumassess))] <- NA

sumassess$Landings <- c(rowSums(lcaton_CURRENT), "2025" = NA)
sumassess$DiscardsDead <- c(rowSums(ddcaton_CURRENT), "2025" = NA)
sumassess$DiscardsSurviving <- c(sCatchEst, NA)

## Format table appropriately
colnames(sumassess) <- c("Year", "R_age1", "RLow", "RHigh", "SSB", "SSBLow", "SSBHigh", "Fbar", "FLow", "FHigh", "Landings", "Discards_Dead", "Discards_Survived")
sumassess <- sumassess[, c("Year", "RLow", "R_age1", "RHigh", "SSBLow", "SSB","SSBHigh", "Landings", "Discards_Dead", "Discards_Survived", "FLow", "Fbar", "FHigh")]

sumassess <- apply_rounding(sumassess)

kable(sumassess,
      caption = paste0("Estimated recruitment, spawning stock biomass (SSB), Landings, Discards and fishing mortality for ages 3-6 of ple.27.21-23.")) 

```







