---
title: "Ple.27.3a.21-32 Annual Assessment Model Interrogation"
author:
  - name: "Elliot J. Brown"
  - name: "Sven Stötera"
  - name: "Casper W. Berg"
date: "`r format(Sys.Date(), '%d/%m/%Y')`"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    code_folding: hide
    code_download: true
---

<style type="text/css">
.main-container {
  max-width: 90% !important;
  margin: auto;
}
p.caption {
  font-size: 0.8em;
  font-style: italic;
}
</style>

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=TRUE,
                      eval = TRUE,
                      warning = FALSE,
                      fig.align = "center",
                      tidy = 'formatR',
                      out.width = "90%")

#===
# Manual settings for figure formatting
#====
# ebpal <- c("#2F3EEA", "#1FD082", "#030F4F", "#F6D04D", "#FC7634", "#F7BBB1", "#E83F48", "#008835", "#79238E")
ebpal <- c("#800080", "#006400", "#D55E00", "#0072B2", "#F0E442", "#009E73", "#E69F00", "#56B4E9", "#CC79A7", "#5DA5DA", "#FF8000", "#89CFF0", "#A52A2A", "#77DD77", "#FFFACD")
#=====
```

# Introduction
## About this document
This interactive document downloads, refits, visualises and interrogates the assessment model from stockassessment.org.  The purpose is to provide an thorough quality control check on the model's performance, and to ensure the working group's (and any other interested parties') understanding of stock dynamics and model estimations. 

Data compilation is done externally, and this document deals only with the finalised input files. All models are build using the State-Space Assessment Model (SAM), and in this version of the document, all model input files and configurations are downloaded directly from the publicly available versions on [stockassessment.org](https://stockassessment.org).

The script that generates this document is one in a series which contribute to the production of advice for this plaice stock. To download the entire series, and associated directory structure, the project is available at GitHub: https://github.com/e-j-brown/Ple.27.21_32_Public.

Most of the code used to process data and create visualisations are hidden in this HTML.  To view each chunk, simply click on the button to the right of the relevant section.  To download the entire R-markdown (.rmd) file including all text and code chunks, use the _code_ button at the top of the page, with the dropdown arrow.

## Preparation
To run the scripts embedded in this document (if you would like to run the assessment yourself) you require the following packages and functions:

```{r message=FALSE, results="hide"}
require(data.table)
require(reshape2)
require(knitr)
require(scales)
require(bookdown)
require(ggplot2)
require(ggthemes)
require(stringr)
require(lattice)
require(plyr)
require(plotly)
require(icesAdvice)
require("stockassessment") # available from https://github.com/fishfollower/SAM
require(parallel)
require(sf)

myprop <- function(df, nomo = "nomo", NoAtLngt = "CANoAtLngt"){
  t <- sum(df$nomo)/sum(df$CANoAtLngt)
}

catchFrac <- function(x, nm, w, frac){
  F <- getF(x)
  Z <- F+nm
  N <- getN(x)
  C <- F/Z*(1-exp(-Z))*N
  return(sum(frac*w*C))
}
```

We also need to set a few variables manually.  
The first is the *DataYear*, which is the year immediately preceeding the assessment year or the year for which the most recent data is available.  
The second is the catch we (ICES) advised as TAC for the last assessment *LastIcesAdvice*, in tonnes. __NOTE__ this value is taken from the advice sheet, to compare changes in advice, not changes in advice relative to actual Allowable Catches.
The third is whether this run is exploratory, or to be used to generate the final advice. This option can be used to determine our conclusions and in some cases, the number of iterations used in parameter estimations. 

We also utilise some reference points to contextualise the output estimates of the model and aid in our understanding of stock status, so we bring those in here, based on the 2024 Benchmark workshop results.

Finally, we set a few paths to be able to find and save stuff.

```{r prepFixedVariables}
DataYear <- 2024
LastIcesAdvice <- 25365 # Advised catch for areas 21-32 (from stock splitting table in advice)
isFinal <- FALSE
# isFinal <- TRUE

## Reference points
MSYBtrigger <- 13460
Bpa <- 13460
Blim <- 11119
SSB_5pcMSY <- 11251.37
Fpa <- 0.149
Fmsy <- 0.149
Fmsy_lower <- 0.137
Fmsy_upper <- 0.149
Fmsy_unconstrained <- 0.46

Flim <- 0.57  # Not based on constraint rule and not used in advice but just FYI

## repository paths
readPath <- paste0("../", (DataYear+1), "/Surveys/")
dataPath <- paste0("DataProcessed/Surveys/")
figPath <- paste0("Figures/")
```

# Baseline

The baseline model we're using to comare this year's model against, is the final model from last year's Benchmark workshop, using catch data up until 2023 and survey data up to Q1 2025. This baseline model is called [ple.27.21-32_BASELAICE_2024_nmG5Fb_mo1_dsmed6](https://stockassessment.org/setStock.php?stock=ple.27.21-32_BASELAICE_2024_nmG5Fb_mo1_dsmed6) and can be found on [stockassessment.org](https://stockassessment.org).

```{r BASEFitFromWeb}
fit_BASE <- fitfromweb("ple.27.21-32_BASELAICE_2024_nmG5Fb_mo1_dsmed6")
```

```{r dataframeSummary_BASE, results='hide', fig.show='hide'}
#===
# Create dataframe of current year's fit for plotting
#====
asum_BASE <- as.data.frame(summary(fit_BASE))
asum_BASE$Year <- as.integer(row.names(summary(fit_BASE)))
ct_BASE <- as.data.frame(catchtable(fit_BASE, obs.show = TRUE))
ct_BASE$Year <- as.integer(rownames(ct_BASE))
rownames(ct_BASE) <- NULL
ct_BASE <- rbind(ct_BASE, data.frame(Estimate = NA, Low=NA, High=NA, sop.catch=NA, Year=as.integer(2024)))
asum_BASE <- merge(x = asum_BASE, y = ct_BASE, by = "Year")
colnames(asum_BASE) <- c("Year", "R_age1", "Rlow", "Rhigh", "SSB", "SSBlow", "SSBhigh", "Fbar", "Flow", "Fhigh", "CatchEst", "Catchlow", "Catchhigh", "CatchObs")
#=====
```

# Current Assessment Model

## Fit Model

This year's assessment model is based on the new Stock Annex procedure from the 2024 Benchmark workshop. This model is also available from [stockassessment.org](https://stockassessment.org), and is called [ple.27.21.32_WGBFAS_CURRENT](https://stockassessment.org/setStock.php?stock=ple.27.21.32_WGBFAS_CURRENT).

```{r 2025_SAFitFromWeb}
fit_CURRENT <- fitfromweb("ple.27.21.32_WGBFAS_2025_SA")
```

```{r dataframeSummary_CURRENT, results='hide', fig.show='hide'}
#===
# Create dataframe of current year's fit for plotting
#====
asum_CURRENT <- as.data.frame(summary(fit_CURRENT))
asum_CURRENT$Year <- as.integer(row.names(summary(fit_CURRENT)))
ct_CURRENT <- as.data.frame(catchtable(fit_CURRENT, obs.show = TRUE))
ct_CURRENT$Year <- as.integer(rownames(ct_CURRENT))
rownames(ct_CURRENT) <- NULL
ct_CURRENT <- rbind(ct_CURRENT, data.frame(Estimate = NA, Low=NA, High=NA, sop.catch=NA, Year=as.integer(DataYear+1)))
asum_CURRENT <- merge(x = asum_CURRENT, y = ct_CURRENT, by = "Year")
colnames(asum_CURRENT) <- c("Year", "R_age1", "Rlow", "Rhigh", "SSB", "SSBlow", "SSBhigh", "Fbar", "Flow", "Fhigh", "CatchEst", "Catchlow", "Catchhigh", "CatchObs")
#=====
```

# Interrogate Model Fits
## Fits to Data

When the model is run we can easily see from warnings/errors if there is a convergence issue.  However, we can also confirm this, explicitly:

- The final model gradient: `r fit_CURRENT$opt$evaluations[2]`
- That there is a positive definite hessian: `r all(eigen(fit_CURRENT$opt$he)$values >0)`

Furthermore, SAM does not utilise "bounds" when fitting the model, and therefore, as standard check of whether model parameters are approaching their bounds is irrelevant.

```{r fig.cap="This year's, stock annex based assessment model's fit to catch data by age (values on the model link function scale).", fig.height=8}
## Extract data from model object
logObs_CURRENT <- split(fit_CURRENT$data$logobs, ceiling(seq_along(fit_CURRENT$data$logobs)/fit_CURRENT$data$maxAgePerFleet[1]))
logPred_CURRENT <- split(fit_CURRENT$rep$predObs, ceiling(seq_along(fit_CURRENT$rep$predObs)/fit_CURRENT$data$maxAgePerFleet[1]))

## Transform data to useable dataframes
### Initialize an empty data frame for observations
logObs_CURRENT_df <- data.frame(age = integer(),
                                         fleet = integer(),
                                         year = integer(),
                                         logObs = numeric())

### Populate the data frame for observations
for (i in seq_along(logObs_CURRENT)) {
  # Calculate the group and year
  year <- ((i - 1) %/% 3) + 1
  fleet <- ((i - 1) %% 3) + 1
  
  ### Create a temporary data frame and append to the main data frame for observations
  temp_df <- data.frame(age = 1:7,
                        fleet = fleet,
                        year = year+2001,
                        logObs = logObs_CURRENT[[i]])
  logObs_CURRENT_df <- rbind(logObs_CURRENT_df, temp_df)
}

### Initialize an empty data frame for predictions
logPred_CURRENT_df <- data.frame(age = integer(),
                                          fleet = integer(),
                                          year = integer(),
                                          logPred = numeric())

### Populate the data frame for predictions
for (i in seq_along(logPred_CURRENT)) {
  # Calculate the group and year
  year <- ((i - 1) %/% 3) + 1
  fleet <- ((i - 1) %% 3) + 1
  
  ### Create a temporary data frame and append to the main data frame for predictions
  temp_df <- data.frame(age = 1:7,
                        fleet = fleet,
                        year = year+2001,
                        logPred = logPred_CURRENT[[i]])
  logPred_CURRENT_df <- rbind(logPred_CURRENT_df, temp_df)
}

## Plot
logPred_CURRENT_df$age <- as.character(logPred_CURRENT_df$age)
logObs_CURRENT_df$age <- as.character(logObs_CURRENT_df$age)

ggplotly(ggplot()+
           geom_point(data = logObs_CURRENT_df[logObs_CURRENT_df$fleet == 1 & logObs_CURRENT_df$year != (DataYear+1), ],
                      mapping = aes(x = year,
                                    y = logObs,
                                    colour = age),
                      shape = 21) +
           geom_line(data = logPred_CURRENT_df[logPred_CURRENT_df$fleet == 1 & logPred_CURRENT_df$year != (DataYear+1), ],
                     mapping = aes(x = year,
                                   y = logPred,
                                   colour = age)) +
           facet_wrap(facets = "age") + #, scales = "free_y") +
           scale_color_manual(values = ebpal, guide = FALSE) +
           guides(colour=FALSE, shape=FALSE) +
           theme_few())

# fitplot(fit_CURRENT, fleets=1)
```


```{r fig.cap="This year's, stock annex based assessment model's fit to Q1 survey data by age (values on the model link function scale).", fig.height=8}
ggplotly(ggplot()+
           geom_point(data = logObs_CURRENT_df[logObs_CURRENT_df$fleet == 2 & logObs_CURRENT_df$year != (DataYear+1), ],
                      mapping = aes(x = year,
                                    y = logObs,
                                    colour = age),
                      shape = 21) +
           geom_line(data = logPred_CURRENT_df[logPred_CURRENT_df$fleet == 2 & logPred_CURRENT_df$year != (DataYear+1), ],
                     mapping = aes(x = year,
                                   y = logPred,
                                   colour = age)) +
           facet_wrap(facets = "age") + #, scales = "free_y") +
           scale_color_manual(values = ebpal, guide = FALSE) +
           guides(colour=FALSE, shape=FALSE) +
           theme_few())
# fitplot(fit_CURRENT, fleets=2)
```

```{r fig.cap="This year's, stock annex based assessment model's fit to Q3/4 survey data by age (values on the model link function scale).", fig.height=8}
ggplotly(ggplot()+
           geom_point(data = logObs_CURRENT_df[logObs_CURRENT_df$fleet == 3 & logObs_CURRENT_df$year != (DataYear+1), ],
                      mapping = aes(x = year,
                                    y = logObs,
                                    colour = age),
                      shape = 21) +
           geom_line(data = logPred_CURRENT_df[logPred_CURRENT_df$fleet == 3 & logPred_CURRENT_df$year != (DataYear+1), ],
                     mapping = aes(x = year,
                                   y = logPred,
                                   colour = age)) +
           facet_wrap(facets = "age") + #, scales = "free_y") +
           scale_color_manual(values = ebpal, guide = FALSE) +
           guides(colour=FALSE, shape=FALSE) +
           theme_few())

# fitplot(fit_CURRENT, fleets=3)
```

```{r fig.cap="Standard Devations by Fleet; medium discard survival estimates.", height=8}
sdplot(fit_CURRENT, marg = c(5,4,1,1))
```

## Residuals
```{r calculateResiduals_CURRENT, warning=FALSE, results='hide'}
resid_CURRENT <- residuals(fit_CURRENT)
resid_CURRENT_df <- data.frame(year = resid_CURRENT$year,
                                        fleet = resid_CURRENT$fleet,
                                        age = resid_CURRENT$age,
                                        observation = resid_CURRENT$observation,
                                        mean = resid_CURRENT$mean,
                                        residual = resid_CURRENT$residual)
resid_CURRENT_df$fleetName <- ifelse(resid_CURRENT_df$fleet == 1, attributes(resid_CURRENT)$fleetNames[1], 
                                              ifelse(resid_CURRENT_df$fleet == 2, attributes(resid_CURRENT)$fleetNames[2], 
                                                     ifelse(resid_CURRENT_df$fleet == 3, attributes(resid_CURRENT)$fleetNames[3], NA)))

resid_CURRENT_df$fleetAltName <- ifelse(resid_CURRENT_df$fleet == 1, "Residual Fishing", 
                                                 ifelse(resid_CURRENT_df$fleet == 2, "Q1 Surveys", 
                                                        ifelse(resid_CURRENT_df$fleet == 3, "Q3/4 Surveys", NA)))
```

```{r fig.cap="Estimated correlations in residual variation between ages for each of the fishing fleet (top) and the two surveys (middle & bottom), from the Model with medium discard mortality."}

if(!all(fit_CURRENT$conf$obsCorStruct=="ID")){ 
  corplot(fit_CURRENT)			  
  # setcap("Estimated correlations", "Estimates correlations between age groups for each fleet")
  # stampit(fit)
} else {
  print("No correlation structure configured for residuals across age.")
}

```

```{r fig.cap="One observation ahead residuals for the three fleets (red/pink = observation higher than model estimate, blue = observation lower than model estimate, size = magnitude of residual).", fig.height=8}
ggplotly(ggplot(resid_CURRENT_df) +
           geom_point(mapping = aes(x = year,
                                    y = age,
                                    size = abs(residual),
                                    colour = residual >= 0),
                      alpha = 0.7) +
           facet_grid(rows = "fleetAltName") +
           scale_colour_manual(values = c("TRUE" = ebpal[8], "FALSE" = ebpal[9])) +
           scale_y_continuous(limits = c(-1, 9), breaks = c(1:9)) +
           guides(colour=FALSE) +
           theme_few())

```

## Jittering
We can also test to see if the model is converging on some local minimum (i.e. it's fitting to some solution close to initialising values that represents noise and not the global solution, that is the proper solution).  To do this, we add random noise to the initial parameter values to see if the model will converge on a different minimum in it's data-space. The easiest way to investigate this is to see if the model fits vary alot depending on the starting values:
```{r Jitter_CURRENT}
jit_CURRENT <- jit(fit = fit_CURRENT)

mt <- as.data.frame(modeltable(jit_CURRENT))
mt$model <- rownames(mt)
rownames(mt) <- NULL

kable(x = mt,
      digits = 3,
      caption = "Measures of  fit for a series of model refits with jitter applied to the input parameters.")
```

# Model Estimates

Now that we've looked at how the model is fitting and performing, we can consider how the estimates of the status look and compare them to current reference points.

```{r recreateAsum, eval=FALSE}
#===
# Create dataframe of current year's fit for plotting
#====
asum_CURRENT <- as.data.frame(summary(fit_CURRENT))
asum_CURRENT$Year <- as.integer(row.names(summary(fit_CURRENT)))
ct_CURRENT <- as.data.frame(catchtable(fit_CURRENT, obs.show = TRUE))
ct_CURRENT$Year <- as.integer(rownames(ct_CURRENT))
rownames(ct_CURRENT) <- NULL
ct_CURRENT <- rbind(ct_CURRENT, data.frame(Estimate = NA, Low=NA, High=NA, sop.catch=NA, Year=as.integer(DataYear+1)))
asum_CURRENT <- merge(x = asum_CURRENT, y = ct_CURRENT, by = "Year")
colnames(asum_CURRENT) <- c("Year", "R_age1", "Rlow", "Rhigh", "SSB", "SSBlow", "SSBhigh", "Fbar", "Flow", "Fhigh", "CatchEst", "Catchlow", "Catchhigh", "CatchObs")
#=====
```


```{r fig.cap= "Estimated spawning stock biomass for ple.27.21-32 and 95% confidence intervals (tonnes). Purple is the baseline model, while green is the outcome of the current stock annex procedure with the latest data. Reference points MSY B-trigger (orange), Bpa (orange, dotted),  and Blim (darker orange) are the horizontal lines."}
pssb <- ggplot() +
  geom_line(data = asum_BASE,
            mapping = aes(x=Year,
                          y=SSB,
                          series="Baseline"),
            colour = ebpal[1]) +
  geom_ribbon(data = asum_BASE,
              mapping = aes(x=Year,
                            ymin=SSBlow,
                            ymax=SSBhigh),
              fill = ebpal[1],
              alpha = 0.3) +
  geom_line(data = asum_CURRENT,
            mapping = aes(x=Year,
                          y=SSB,
                          series="Current Year SA"),
            colour = ebpal[2]) +
  geom_ribbon(data = asum_CURRENT,
              mapping = aes(x=Year,
                            ymin=SSBlow,
                            ymax=SSBhigh),
              fill = ebpal[2],
              alpha = 0.3) +
  geom_hline(yintercept = Bpa,
             colour = ebpal[3],
             linetype = "dotted",
             aes(series="Bpa")) +
  geom_hline(yintercept = MSYBtrigger,
             colour = ebpal[11],
             mapping = aes(series="MSY B-trigger")) +
  geom_hline(yintercept = Blim,
             colour = ebpal[3],
             mapping = aes(series="Blim")) +
  theme_clean()

ggplotly(pssb, tooltip = c("series", "x", "y"))
```


```{r fig.cap= "Annual fishing mortality estimates for ple.27.21-32 ages 3-5 and point wise 95% confidence intervals are shown by line and shaded area.   Purple is the baseline model, while green is the outcome of the current stock annex procedure with the latest data. Reference points Fmsy (orange), Fpa (orange, dotted), and Flim (darker orange) are the horizontal lines."}
pfbar <- ggplot() +
  geom_line(data = asum_BASE,
            mapping = aes(x=Year,
                          y=Fbar,
                          series="Baseline"),
            colour = ebpal[1]) +
  geom_ribbon(data = asum_BASE,
              mapping = aes(x=Year,
                            ymin=Flow,
                            ymax=Fhigh,
                            series="Baseline"),
              fill = ebpal[1],
              alpha = 0.3) +
  geom_line(data = asum_CURRENT,
            mapping = aes(x=Year,
                          y=Fbar,
                          series="Current Year SA"),
            colour = ebpal[2]) +
  geom_ribbon(data = asum_CURRENT,
              mapping = aes(x=Year,
                            ymin=Flow,
                            ymax=Fhigh,
                            series="Current Year SA"),
              fill = ebpal[2],
              alpha = 0.3) +
  geom_hline(yintercept = Fpa,
             colour = ebpal[3],
             linetype = "dotted",
             aes(series="Fpa")) +
  geom_hline(yintercept = Fmsy,
             colour = ebpal[11],
             aes(series="Fmsy")) +
  geom_hline(yintercept = Flim,
             colour = ebpal[3],
             aes(series="Flim")) +
  theme_clean()

ggplotly(pfbar, tooltip = c("series", "x", "y"))
```

```{r fig.cap= "Annual recruitment estimates for ple.27.21-32 and point wise 95% confidence intervals are shown by line and shaded area (numbers). Purple is the baseline model, while green is the outcome of the current stock annex procedure with the latest data."}
prec <- ggplot() +
  geom_line(data = asum_BASE,
            mapping = aes(x=Year,
                          y=R_age1,
                          series="Baseline"),
            colour = ebpal[1]) +
  geom_ribbon(data = asum_BASE,
              mapping = aes(x=Year,
                            ymin=Rlow,
                            ymax=Rhigh,
                            series="Baseline"),
              fill = ebpal[1],
              alpha = 0.3) +
  geom_line(data = asum_CURRENT,
            mapping = aes(x=Year,
                          y=R_age1,
                          series="Current Year SA"),
            colour = ebpal[2]) +
  geom_ribbon(data = asum_CURRENT,
              mapping = aes(x=Year,
                            ymin=Rlow,
                            ymax=Rhigh,
                            series="Current Year SA"),
              fill = ebpal[2],
              alpha = 0.3) +
  theme_clean()

ggplotly(prec, tooltip = c("series", "x", "y"))
```

```{r fig.cap= "Annual catch estimates and 95% confidence intervals (line and shaded area, respectively) for ple.27.21-32 and point annual observations (points).  Purple is the baseline model, while green is the outcome of the current stock annex procedure with the latest data, and orange are the observations."}
pcat <- ggplot() +
  geom_line(data = asum_BASE[asum_BASE$Year != (DataYear), ],
            mapping = aes(x=Year,
                          y=CatchEst,
                          series="Baseline"),
            colour = ebpal[1]) +
  geom_ribbon(data = asum_BASE[asum_BASE$Year != (DataYear), ],
              mapping = aes(x=Year,
                            ymin=Catchlow,
                            ymax=Catchhigh,
                            series="Baseline"),
              fill = ebpal[1],
              alpha = 0.3) +
  geom_line(data = asum_CURRENT[asum_CURRENT$Year != (DataYear+1), ],
            mapping = aes(x=Year,
                          y=CatchEst,
                          series="Current Year SA"),
            colour = ebpal[2]) +
  geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != (DataYear+1), ],
              mapping = aes(x=Year,
                            ymin=Catchlow,
                            ymax=Catchhigh,
                            series="Current Year SA"),
              fill = ebpal[2],
              alpha = 0.3) +
  geom_point(data = asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ],
             mapping = aes(x=Year,
                           y=CatchObs,
                           series="Observed Catches"),
             shape = 3,
             colour = ebpal[3],
             fill = ebpal[3]) +
  theme_clean()

ggplotly(pcat, tooltip = c("series", "x", "y"))
```

# Model Performance and Sensitivities
## Leave-One-Out Analyses
A leave-one-out analysis is a form of sensitivity analysis, showing the impact the data from each tuning fleet has on the estimation of the key variables being estimated; namely SSB, F and recruitment.  

First we must run the leave-one-out analysis which refits the model in two iterations, removing one survey at a time. Then we can plot each of these new model fits over the full model to see the impact the removal of each has. 

```{r LeaveOneOut_CURRENT}
LO_CURRENT <- leaveout(fit_CURRENT)

#=== 
# Get data from sam objects and generate useable dataframes
#====
q1mat <- as.data.frame(summary(LO_CURRENT$`w.o. Q1IBTS+BITS1-7+`))
q3mat <- as.data.frame(summary(LO_CURRENT$`w.o. Q4IBTS+BITS1-7+`))


woq1 <- data.frame(Year = as.integer(row.names(summary(LO_CURRENT$`w.o. Q1IBTS+BITS1-7+`))),
                   SSB = q1mat$SSB,
                   Fbar = q1mat$`Fbar(3-6)`,
                   R_age1 = q1mat$`R(age 1)`,
                   CatchEst = catchtable(LO_CURRENT$`w.o. Q1IBTS+BITS1-7+`)[,1],
                   series = rep("wo_Q1", times = nrow(q1mat)))

woq34 <- data.frame(Year = as.integer(row.names(summary(LO_CURRENT$`w.o. Q4IBTS+BITS1-7+`))),
                    SSB = q3mat[, "SSB"],
                    Fbar = q3mat[,"Fbar(3-6)"],
                    R_age1 = q3mat[,"R(age 1)"],
                    CatchEst = c(catchtable(LO_CURRENT$`w.o. Q4IBTS+BITS1-7+`)[,1], setNames(NA, (DataYear+1))),
                    series = rep("wo_Q34", times = (nrow(q3mat))))


asum_CURRENT$series <- rep("full", times = nrow(asum_CURRENT))
asumi_CURRENT <- asum_CURRENT[, c("Year", "SSB", "Fbar", "R_age1", "CatchEst" ,"series")]
#=====

losum_CURRENT <- rbind(woq1, woq34, asumi_CURRENT)
losum_CURRENT[losum_CURRENT$Year == (DataYear+1), "Fbar"] <- NA  # Remove estimate for assessment year as it makes no sense to estimate without catch data.

```

```{r fig.cap="Leave-one-out re-fits for the model with medium discard survival (without Q1 survey = blue, without Q3/4 survey = purple), overlain with full model estimates (black line and grey ribbon) of SSB (top left), F (top right), Recruitment (bottom left), and catch (bottom right; observations as yellow + and ribbon broken by interactive tool)."}
# ssbplot(LO)
lossb_CURRENT <- ggplotly(ggplot() +
                                     geom_line(data = losum_CURRENT,
                                               mapping = aes(x=Year,
                                                             y=SSB,
                                                             colour = series)) +
                                     geom_ribbon(data = asum_CURRENT,
                                                 mapping = aes(x=Year,
                                                               ymin=SSBlow,
                                                               ymax=SSBhigh),
                                                 fill = "black",
                                                 alpha = 0.2) +
                                     scale_colour_manual(values = c("black", ebpal[8:9])) +
                                     theme_clean() +
                                     theme(axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95)))

# Fplot(LO)
lof_CURRENT <- ggplotly(ggplot() +
                                   geom_line(data = losum_CURRENT,
                                             mapping = aes(x=Year,
                                                           y=Fbar,
                                                           colour = series)) +
                                   geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != (DataYear+1),],
                                               mapping = aes(x=Year,
                                                             ymin=Flow,
                                                             ymax=Fhigh),
                                               fill = "black",
                                               alpha = 0.3) +
                                   scale_colour_manual(values = c("black", ebpal[8:9])) +
                                   theme_clean() +
                                   theme(axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95)))

# recplot(LO)
lorec_CURRENT <- ggplotly(ggplot() +
                                     geom_line(data = losum_CURRENT,
                                               mapping = aes(x=Year,
                                                             y=R_age1,
                                                             colour = series)) +
                                     geom_ribbon(data = asum_CURRENT,
                                                 mapping = aes(x=Year,
                                                               ymin=Rlow,
                                                               ymax=Rhigh),
                                                 fill = "black",
                                                 alpha = 0.3) +
                                     scale_colour_manual(values = c("black", ebpal[8:9])) +
                                     theme_clean() +
                                     theme(axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95)))

# Catch plot (LO)
loca_CURRENT <- ggplotly(ggplot() +
                                    geom_line(data = losum_CURRENT,
                                              mapping = aes(x=Year,
                                                            y=CatchEst,
                                                            colour = series))+
                                    geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != (DataYear+1), ],
                                                mapping = aes(x=Year,
                                                              ymin=Catchlow,
                                                              ymax=Catchhigh),
                                                fill = "black",
                                                alpha = 0.3) +
                                    geom_point(data = asum_CURRENT[asum_CURRENT$Year != (DataYear+1), ],
                                               mapping = aes(x=Year,
                                                             y=CatchObs),
                                               shape = 3,
                                               colour = ebpal[5],
                                               fill = ebpal[5]) +
                                    ylab("Catch (tonnes)") +
                                    scale_colour_manual(values = c("black", ebpal[8:9]))+ 
                                    theme_clean() +
                                    theme(axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95)))

layout(subplot(lossb_CURRENT, lof_CURRENT, lorec_CURRENT, loca_CURRENT, nrows = 2, shareX = TRUE, titleY = TRUE), showlegend=FALSE)
```

### Retrospectives
```{r calculateRetros_CURRENT}
RETRO_CURRENT<-retro(fit_CURRENT, year=5)
rho_CURRENT <- mohn(RETRO_CURRENT, lag = 1)

## Make RETROs in better plotting format
ret_CURRENT_df <- asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ]

for(i in 1:length(RETRO_CURRENT)){
  tsum <- as.data.frame(summary(RETRO_CURRENT[[i]]))
  tsum$Year <- as.integer(row.names(summary(RETRO_CURRENT[[i]])))
  tsum <- cbind(tsum, catchtable(RETRO_CURRENT[[i]], obs.show = TRUE))
  tsum$series <- as.character(rep(i, nrow(tsum)))
  colnames(tsum) <- c("R_age1", "Rlow", "Rhigh", "SSB", "SSBlow", "SSBhigh", "Fbar", "Flow", "Fhigh", "Year", "CatchEst", "Catchlow", "Catchhigh", "CatchObs", "series")
  tsum <- tsum[tsum$Year != max(tsum$Year), ]
  ret_CURRENT_df <- rbind(ret_CURRENT_df, tsum)
}
ret_CURRENT_df$series <- factor(ret_CURRENT_df$series, levels = c("full", "1", "2", "3", "4", "5"))
```

```{r fig.cap="Retrospective analyses for SSB (top-left), F for ages 3-5 (top-right), recruitment (bottom-left), and catch (bottom-right), from the model with medium discard survival."}
# ssbplot(RETRO)
retssb_CURRENT <- layout(ggplotly(ggplot() +
                                             geom_line(data = ret_CURRENT_df,
                                                       mapping = aes(x=Year,
                                                                     y=SSB,
                                                                     colour = series)) +
                                             geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ],
                                                         mapping = aes(x=Year,
                                                                       ymin=SSBlow,
                                                                       ymax=SSBhigh),
                                                         fill = "black",
                                                         alpha = 0.2) +
                                             annotate(geom = "text",
                                                      y = max(ret_CURRENT_df$SSBhigh)*0.85,
                                                      x = ((max(ret_CURRENT_df$Year)-min(ret_CURRENT_df$Year))*0.20)+min(ret_CURRENT_df$Year),
                                                      label = paste0("Mohn's Rho = ", round(rho_CURRENT[2], digits = 3))) +
                                             scale_colour_manual(values = c("black", ebpal[(length(ebpal)-4):length(ebpal)])) +
                                             theme_clean() +
                                             theme(legend.position = "none",
                                                   axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95))),
                                  showlegend = FALSE)

# Fplot(RETRO)
retf_CURRENT <- layout(ggplotly(ggplot() +
                                           geom_line(data = ret_CURRENT_df,
                                                     mapping = aes(x=Year,
                                                                   y=Fbar,
                                                                   colour = series)) +
                                           geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ],
                                                       mapping = aes(x=Year,
                                                                     ymin=Flow,
                                                                     ymax=Fhigh),
                                                       fill = "black",
                                                       alpha = 0.3) +
                                           annotate(geom = "text",
                                                    y = max(ret_CURRENT_df$Fhigh)*0.85,
                                                    x = ((max(ret_CURRENT_df$Year)-min(ret_CURRENT_df$Year))*0.80)+min(ret_CURRENT_df$Year),
                                                    label = paste0("Mohn's Rho = ", round(rho_CURRENT[3], digits = 3))) +
                                           scale_colour_manual(values = c("black", ebpal[(length(ebpal)-4):length(ebpal)])) +
                                           theme_clean() +
                                           theme(legend.position = "none",
                                                 axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95))),
                                showlegend = FALSE)

# recplot(RETRO)
retrec_CURRENT <- layout(ggplotly(ggplot() +
                                             geom_line(data = ret_CURRENT_df,
                                                       mapping = aes(x=Year,
                                                                     y=R_age1,
                                                                     colour = series)) +
                                             geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ],
                                                         mapping = aes(x=Year,
                                                                       ymin=Rlow,
                                                                       ymax=Rhigh),
                                                         fill = "black",
                                                         alpha = 0.3) +
                                             annotate(geom = "text",
                                                      y= max(ret_CURRENT_df$Rhigh)*0.85,
                                                      x = ((max(ret_CURRENT_df$Year)-min(ret_CURRENT_df$Year))*0.20)+min(ret_CURRENT_df$Year),
                                                      label = paste0("Mohn's Rho = ", round(rho_CURRENT[1], digits = 3))) +
                                             scale_colour_manual(values = c("black", ebpal[(length(ebpal)-4):length(ebpal)])) +
                                             theme_clean() +
                                             theme(legend.position = "none",
                                                   axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95))),
                                  showlegend = FALSE)

## Catch plot (RETRO)
retca_CURRENT <- ggplotly(ggplot() +
                                     geom_line(data = ret_CURRENT_df,
                                               mapping = aes(x=Year,
                                                             y=CatchEst,
                                                             colour = series))+
                                     geom_ribbon(data = asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ],
                                                 mapping = aes(x=Year,
                                                               ymin=Catchlow,
                                                               ymax=Catchhigh),
                                                 fill = "black",
                                                 alpha = 0.3) +
                                     geom_point(data = asum_CURRENT[asum_CURRENT$Year != max(asum_CURRENT$Year), ],
                                                mapping = aes(x=Year,
                                                              y=CatchObs),
                                                shape = 3,
                                                colour = ebpal[5],
                                                fill = ebpal[5]) +
                                     ylab("Catch (tonnes)") +
                                     scale_colour_manual(values = c("black", ebpal[(length(ebpal)-4):length(ebpal)]))+ 
                                     theme_clean() +
                                     theme(axis.title.y.left = element_text(vjust = -0.05, hjust = 0.95)))

ret_fp <- style(subplot(retssb_CURRENT, retf_CURRENT, retrec_CURRENT, retca_CURRENT, nrows = 2, shareX = FALSE, shareY = FALSE, titleY = TRUE),
                showlegend=FALSE,
                traces = c(8:((7*4)+2)))

layout(ret_fp, legend = list(orientation = 'h', x = 0.5, xanchor = 'center', y = -0.05, yanchor = 'top', borderwidth = 0))
```




